Decomposition-based Job-shop Scheduling with Constrained Clustering

### 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
Features

\begin{document}
%
\title{Decomposition-based Job-shop Scheduling with Constrained Clustering}%\thanks{Supported by organization x.}}
%
\titlerunning{Decomposition-based JSP with Constrained Clustering}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958} \and
Martin Gebser\inst{1,2}\orcidID{0000-0002-8010-4752}}
%
\authorrunning{M. El-Kholany et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Alpen-Adria-Universität Klagenfurt, % Klagenfurt, 
Austria\\% 
% \email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}%\\
%\url{http://www.springer.com/gp/computer-science/lncs}
\and
Technische Universität Graz, % Graz, 
Austria\\%
% \email{mgebser@ist.tugraz.at}
\email{firstname.lastname@aau.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Scheduling is a crucial problem appearing in various domains, such as manufacturing, transportation, or healthcare, where % . In most problem definitions, 
the goal is to schedule % a given set of 
given operations on available resources such that the operations are completed as soon as possible. Unfortunately, most scheduling problems cannot be solved efficiently, % Therefore, the 
so that research on suitable approximation methods is of primary importance.
%
This work introduces a novel approximation approach based on problem decomposition with data mining methodologies. We propose a constrained clustering algorithm to group operations into clusters, corresponding to time windows in which the operations must be scheduled. 
The decomposition process consists of two main phases. 
First, features are extracted, either from the problem itself or from solutions obtained by heuristic methods, to predict the execution sequence of operations on each resource.
% The first phase is to extract features to predict the sequence of operations on each resource. These features are extracted either from the problem itself or from solutions obtained by other heuristics.
The second phase % is to develop
deploys our constrained clustering algorithm to assign each operation into a time window. We then schedule the operations by time windows using Answer Set Programming. Evaluation results show that our proposed approach outperforms other heuristic schedulers in most cases, where incorporating features like \textit{Remaining Processing Time}, \textit{Machine Load}, and \textit{Earliest Starting Time} significantly improves the solution quality.

\keywords{Job-shop Scheduling  Problem \and Constrained Clustering \and Time Windows \and Answer Set Programming}
\end{abstract}
%
%
%
\section{Introduction}
Scheduling is one of the most crucial problems in various industrial, transportation, or healthcare applications
%\todo{add 1-2 citations for each application}
 \cite{pezzella2008genetic,chaudhry2016research,nouiri2018effective,demirbilek2019dynamically,schoenfelder2020nurse,wang2019routing,janakbhai2021blockchain}.

@@

\section{Features extraction}
\label{sec:features}
Data mining merges with machine learning and statistical techniques to gain information and new knowledge from the databases\cite{ismail2012production,nasiri2019data}. Examples of data mining tasks are classification, clustering, and sequential pattern mining. In this work, we aim to use clustering to study the operations' pattern to find a good sequencing of the operations on each machine that leads us to obtain a near-optimal solution of the JSP. In order to use the clustering approach, we should have some features which describe the characteristics of the problem. These features would aid us in the decision-making process, hence, reaching a near-optimal solution. \\

Some of the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} have extracted the following features: priority, processing time, remaining processing time, machine load, and route position. Most of the researchers converted the attributes into qualitative data. As mentioned in these studies, the main reason for this conversion is to make the decision rules more generic, which can be applied to several scheduling problems. However, in our case, we will not convert the data into classes because we will use these features to measure the similarities/dissimilarities between the operations in the clustering process, which requires having numerical values.
We present the features that we considered:
%\paragraph{Process.}
%The two attributes \textit{Processing Time (PT)}  and \textit{Remain Processing Time (RPT)} are dependent on the problem domain. The \textit{PT} represents the execution time of a particular operation in a machine. However, \textit{RPT} is the aggregate processing time for all subsequent operations of that job. \todo[inline]{we need to provide rationale behind the features, why do we think they are useful for the clusering algorithm. Example of values extracted for the sample instance in Table 1 would also improve the readability of the paper.}
%\todo[inline]{Why in all other paragraphs we are talking only about one feature at a time and here about two. Can we split them? Like \\paragraph\{Processing Time (PT)\} and \\paragraph\{Remain Processing time (RPT)\}}
\begin{description}
  \item[Processing Time (PT)] that represents the execution time of a particular operation in a machine.

  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job.

@@

  \item[Time Length of a Job (TLJ)] is an attribute describing the total processing time needed to execute all operations of a job. 
  %It is the summation of processing time for all operations of that job.

  \item[Earliest Starting Time (EST)] attribute represents the earliest possible execution time for an operation. It is calculated by accumulating the processing time of all preceding operations. This feature has a specific value $\mathit{Zero}$ if the operation has no predecessors.

  \item[Machine Load (ML)] is a property of a machine that shows how much time this machine requires to execute all the operations assigned to it. In the initial state, \textit{ML} corresponds to the total processing time of all operations assigned to that machine. For each step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}.

  \item[Starting Time (ST)] is a set of features representing the starting execution time of each operation by solving the problem using a heuristic. In our work, we consider EST (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr), see \cite{jones1998survey} for an overview.
To compute the starting time, we use a simple greedy algorithm \cite{el2020job}, that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine. In the case of FIFO, the algorithm selects an operation that appeared first in the queue. Similarly, for MTWR, the algorithm selects the operation that belongs to the job with the longest processing time needed to be processed.

  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr). Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and its earliest possible starting time. The latter is the maximum out of the end processing time of the operation's predecessor and the time at which the required machine became available.
\end{description}

The following Table~\ref{tab3} shows a sample of the features that we extracted for the decomposition process. Each row in the table represents an operation, and it contains its value according to each feature. The second column shows the \textit{Remaining Processing Time (RPT)} of a job. For instance, the first job $J_1$ consists of $3$ operations with a total processing time of $17$. In the beginning, the \textit{RPT} of $O_{1,1}$ is $17$ because no operations are executed. After finishing the first operation of job $1$, the \textit{RPT} became $8$ because the processing time of the completed operation $O_{1,1}$ is excluded. When the second operation of job $1$ is completed, the \textit{RPT} of that job is $5$. \\

The third column represents the \textit{Earliest Starting Time} which we described earlier. The last column is the \textit{Machine Load} which is the summation of the processing time for all operations assigned to a machine. For example, Machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$, the load of machine $3$ at time $0$ is 10. We assumed that the machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}. After completing the operation $O_{2,1}$, the machine will target the next operation, which has the smallest value of \textit{EST} and the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$, and so on till completing all the operations assigned to a particular machine.

\begin{table}[h!]
\fontsize{12}{9}
\setlength{\tabcolsep}{10.0pt}
\centering
  \begin{center}
  \caption{Time Window Assignment.}
  \label{tab3}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c  c  c  c}
      \textbf{Operation} & \textbf{RPT} & \textbf{EST} & \textbf{ML}\\
      \hline
										\\
      $O_{1,1}$  & 17 & 0   & 20	\\
      $O_{1,2}$  & 8  & 9   & 3		\\
      $O_{1,3}$  & 5  & 12 & 5		\\
										\\
      $O_{2,1}$  & 12 & 0   & 10	\\
      $O_{2,2}$  & 8  & 4   & 11	\\
      $O_{2,3}$  & 2  & 10 & 7		\\
					    					\\
      $O_{3,1}$  & 12 & 0  & 11	\\
      $O_{3,2}$  & 8  & 4  & 6		\\
      $O_{3,3}$  & 5  & 7  & 5		\\
    \end{tabular}
%\end{adjustwidth}

>>>>

\begin{document}
%
\title{Decomposition-based Job-shop Scheduling with Constrained Clustering}%\thanks{Supported by organization x.}}
%
\titlerunning{Decomposition-based JSP with Constrained Clustering}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Mohammed M. S. El-Kholany\inst{1}\orcidID{0000-0002-1088-2081} \and
Konstantin Schekotihin\inst{1}\orcidID{0000-0002-0286-0958} \and
Martin Gebser\inst{1,2}\orcidID{0000--0000-0000-0000}}
%
\authorrunning{M. El-Kholany et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Alpen-Adria-Universität Klagenfurt, Klagenfurt, Austria 
\email{\{mohammed.el-kholany, konstantin.schekotihin and martin.gebser\}@aau.at}\\
%\url{http://www.springer.com/gp/computer-science/lncs} \and
Technische Universität Graz, Graz, Austria\\
\email{mgebser@ist.tugraz.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Scheduling is a crucial problem appearing in various domains, such as manufacturing, transportation, or healthcare. In most problem definitions, the goal is to schedule a given set of operations on available resources to complete the operations as early as possible. Unfortunately, most scheduling problems cannot be solved efficiently. Therefore, the research of suitable approximation methods is of primary importance.

This work suggests a novel approximation approach based on problem decomposition with data mining methodologies. This study proposes a constrained clustering algorithm to group the operations into clusters corresponding to time windows in which these operations must be scheduled. 
The decomposition process depends on two main phases. The first phase is to extract features to predict the sequence of the operations on each resource. These features are extracted either from the problem itself or from solutions obtained by other heuristics. The second phase is to develop a constrained clustering algorithm to assign each operation into a time window. We solved the problem using the Answer Set Programming. Evaluation results show that our proposed outperformed other heuristic schedulers in most cases, where features, like \textit{Remaining Processing Time}, \textit{Machine Load}, and \textit{Earliest Starting Time}, contributed significantly to the solution quality.

\keywords{Job-shop Scheduling  Problem \and Constrained Clustering \and Time Windows \and Answer Set Programming.}
\end{abstract}
%
%
%
\section{Introduction}
Scheduling is one of the most crucial problems in various industrial, transportation, or healthcare applications
%\todo{add 1-2 citations for each application}
 \cite{pezzella2008genetic,chaudhry2016research,nouiri2018effective,demirbilek2019dynamically,schoenfelder2020nurse,wang2019routing,janakbhai2021blockchain}.

@@

\section{Feature extraction}
\label{sec:features}
Application of machine learning methods to scheduling problems requires careful selection of data describing the hidden dependencies between operations of different jobs \cite{ismail2012production,nasiri2019data}. Clustering methods, which we aim to apply in our approach, are not an exception. Thus, a clustering method requires an informative set of features describing jobs, their operations, and machines of an instance to find patterns resulting in a beneficial decomposition of all operations in time windows.

Methods suggested in the previous work \cite{koonce2000using, harrath2002genetic, shahzad2010discovering, ismail2012production, adibi2014clustering, nasiri2019data} used characterizations of scheduling instances based on the following features: priority, processing time, remaining processing time, machine load, and route position. 
Most of the approaches converted the quantitative values of the attributes into qualitative data. As mentioned in these studies, the main reason for this conversion was to make the heuristic dispatching rules more generic and allow their application to several scheduling problems. 
In this work, we suggest a method that can work directly on the sets of features and does not require any problem-specific transformations. However, our clustering method for JSP instance decomposition requires all features to have numerical values allowing distance measures to estimate similarities/dissimilarities between operations of jobs.
%
Therefore, we suggest the following features:
\begin{description}
  \item[Processing Time (PT)] that represents the execution time of a particular operation in a machine. This feature is provided in every JSP instance, like in the sample instance presented in Table \ref{tab1}.

@@

  \item[Remaining Processing Time (RPT)] attribute corresponds to the total processing time for all subsequent operations of a job. For example, Table~\ref{tab3} shows a set of \textit{RPT} values computed for the sample instance. Since the first job $J_1$ consists of $3$ operations with a total processing time of $17$, \textit{RPT} of $O_{1,1}$ is $17$ because it has no preceding operations. After finishing the first operation of $J_1$, the \textit{RPT} becomes $8$ because the processing time of the completed operation $O_{1,1}$ is subtracted. Similarly, when the second operation of $J_1$ is completed, the \textit{RPT} of the last operation is $5$.

  \item[Time Length of a Job (TLJ)] is an attribute describing the total processing time needed to execute all operations of a job. Since all operations of a job have the same feature value, this feature can only be used to differentiate between jobs.
  %It is the summation of processing time for all operations of that job.

  \item[Earliest Starting Time (EST)] attribute represents the earliest possible execution time for an operation. It is calculated by accumulating the processing time of all preceding operations. This feature has a specific value $\mathit{Zero}$ if the operation has no predecessors.\todo{This value does not appear in the table, why?}

  \item[Machine Load (ML)] is a property of a machine that shows how much time this machine requires to execute all the operations assigned to it. In the initial state, \textit{ML} corresponds to the total processing time of all operations assigned to that machine. In each time step, the load of a machine is decreased by the processing time of an operation assigned to the machine and has the smallest \textit{EST}. For example, machine $3$ has to process operations $\{ O_{1,2}, O_{2,1}, O_{3,2} \}$. The load of machine $3$ at time $0$ is 10. We assumed that the machine would start processing the operation $O_{2,1}$ that has the smallest \textit{EST}. After completing the operation $O_{2,1}$, the machine will target the next operation, which has the smallest value of \textit{EST} and the load of the machine is decreased by the processing time of the completed operation $O_{2,1}$, and so on till completing all the operations assigned to a particular machine.

  \item[Starting Time (ST)] is a set of features representing the starting execution time of each operation by solving the problem using a heuristic. In our work, we consider EST (\stest), First-In-First-Out (\stfifo), and Most Total Work Remaining (\stmtwr), see \cite{jones1998survey} for an overview.
  To compute the starting time, we use a simple greedy algorithm \cite{el2020job} that at each step selects a not scheduled operation according to the heuristics and schedules it to a machine. In the case of FIFO, the algorithm selects an operation that appeared first in the queue. Similarly, for MTWR, the algorithm selects the operation that belongs to the job with the longest processing time from the queue.

  \item[Waiting Time (WT)] is also a set of features, like ST, for which computation we used the same greedy algorithm as above with three different heuristics EST (\wtest), FIFO (\wtfifo), and MTWR (\wtmtwr). Each feature in \textit{WT} set is the difference between the scheduled starting time of a particular operation and its earliest possible starting time. The latter is the maximum out of the end processing time of the operation's predecessor and the earliest time at which the required machine became available.

  \todo{It would be nice to add examples of other features to the table. If this is complicated, then only of \wtest and \stest.}
\end{description}

\begin{table}[h!]
\setlength{\tabcolsep}{10.0pt}
\centering
  \begin{center}
  \caption{Features extracted for the sample problem instance.}
  \label{tab3}
 %\begin{adjustwidth}{0.0cm}{}
    \begin{tabular}{c  c  c  c}
      \textbf{Operation} & \textbf{RPT} & \textbf{EST} & \textbf{ML}\\
      \hline
										\\
      $O_{1,1}$  & 17 & 0   & 20	\\
      $O_{1,2}$  & 8  & 9   & 3		\\
      $O_{1,3}$  & 5  & 12 & 5		\\
										\\
      $O_{2,1}$  & 12 & 0   & 10	\\
      $O_{2,2}$  & 8  & 4   & 11	\\
      $O_{2,3}$  & 2  & 10 & 7		\\
					    					\\
      $O_{3,1}$  & 12 & 0  & 11	\\
      $O_{3,2}$  & 8  & 4  & 6		\\
      $O_{3,3}$  & 5  & 7  & 5		\\
    \end{tabular}
%\end{adjustwidth}
